Efficient Parallelization of $\ell_1$-Minimization for Face Recognition \thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}

### 2010-11-09 paper/main.tex
Removing cruft...
changed two subsection titles
started section 2 and 5

	%Major steps:
	%while:
		%mu *= rho
		%lambda += mu(y - Ax - e)
		%while:
			%e <- shrink(y-Ax+lambda/mu, 1/mu)
			%x <- Apinvt'*(y-e+lamba/mu)
We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
\begin{algorithm}[h]
\caption{\bf (Augmented Lagrange Multiplier Method Used in Alignment Inner Loop)}
\begin{algorithmic}[1]
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
\ENDWHILE
\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{algorithmic}
\label{alg:alm}
\end{algorithm}

@@

\section{Efficient Implementation on the CPU}
% With a reshuffling of the order of operations and a change of the termination
% condition, the inner loop takes the form of two matrix-vector multiplications
% and a sequence of vector-vector operations.
\subsection{CPU Hardware Parallelism}
High-level discussion of CPU parallelism, i.e.
thread-level, vector-level, cache size, cache locality
\subsection{Tuning the GEMV operation}
When matrix-vector operations are called sequentially on the same data,
the operation can be blocked so that each thread operates on the same block of data for each operation.

@@

\section{Efficient Implementation on the CPU}
\subsection{GPU Hardware Parallelism}
High-level description of GPU parallelism, i.e.

@@

Optimization:
\begin{equation}
\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}

>>>>

The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
\begin{equation}
F(x) = f(x) + \lambda g(x).
\end{equation}

Because the optimal lagrange multiplier values are not known, ALM iterates between estimating the solution and the lagrange multipier values.

In order to solve the recognition stage, we consider the $l_1$ problem:
\begin{equation}
\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
which can be formulated as
\begin{equation}
L_u(x,y) = \|x\|_1 + <\y, \b - A\x - \e> + \frac{\mu}2 \| \b-A\x-\e \|_2^2
\end{equation}

where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.

We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the
largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
\begin{algorithm}[h]
\caption{\bf (Augmented Lagrange Multiplier Method Used in Alignment Inner Loop)}
\begin{algorithmic}[1]
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
\ENDWHILE
\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{algorithmic}
\label{alg:alm}
\end{algorithm}

In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
\begin{equation}
\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
which is solved in a similar fashion.

@@

\section{CPU Implementation}
% With a reshuffling of the order of operations and a change of the termination
% condition, the inner loop takes the form of two matrix-vector multiplications
% and a sequence of vector-vector operations.
\subsection{CPU Hardware Parallelism}
High-level discussion of CPU parallelism, i.e.
thread-level, vector-level, cache size, cache locality
\subsection{Tuning the GEMV operation}
When matrix-vector operations are called sequentially on the same data,
the operation can be blocked so that each thread operates on the same block of data for each operation.

@@

\section{GPU Implementation}
\subsection{GPU Hardware Parallelism}
High-level description of GPU parallelism, i.e.

@@

In this section, we benchmark the performance of MATLAB, BLAS MKL, and GPU Cuda implementations of the Primal ALM algorithm on synthetic data to find the point at which the GPU version runs faster than CPU versions.
To simply the benchmark, we cap the number of inner iterations at 50 and outer iterations at 5000 and run the algorithm to convergence.
CPU code was run on an Intel Dual Socket Quad Core with 8 gigs of DDR3 RAM.  GPU code was run on a GTX480
We ran the A matrix with different ratios between m:n from 10:1:5 for m:2000:25:8000 with an x vector of 10 percent sparsity.  Tolerance levels were empirically seen to allow best performance with the inner loop tolerance set at one order of magnitude lower than the outer loop tolerance.

For figures ...

As the dimensions of A increase, so does the significance of the size of AtA.  Until the A and AtA matrix reach x Mb, the CPU version runs faster than the GPU.

Optimization:
\begin{equation}
\min_{\x, \e} \|\x\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
